---
title: "final project"
author: "Eric Tu"
date: "2025-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Introduction:

  A student's GPA is one of the most common forms of academic measurement that many people, such as Companies, Professors, and even student themselves, use to get an idea of one's academic level. GPA is calculated by assigning each letter grade a number of points and then averaging them. There are many uses for this metric, such as helping students assess how well they’re doing, helping companies decide who to hire, and helping university admission officers evaluate applicants. Because of these uses, it’s important to maintain a high GPA.

  Everyone in our group was curious about how students at UW–Madison have kept up with this standard, which led us to a data set containing all course grades at UW–Madison from 2006 to 2017. Something that isn't clearly stated anywhere but is very interesting to know about is how different our GPA is today compared to the past. Since we had access to historical grades, we decided to compare the last semester’s average GPA to the average in our data set. We found the last semester’s average GPA to be 3.492, which gave us a benchmark for comparison. After calculating the overall average GPA from the 2006–2017 data, we discovered that the data set’s average GPA was actually higher than last semester’s reported value. Therefore, our thesis is that the average GPA at UW–Madison has decreased compared to the historical average, indicating a meaningful shift in academic performance over time.


Background:

Statistical Analysis:

null hypothesis: the average gpa at UW-Madison is 3.492
alternate hypothesis: the average gpa at UW-Madison is not equal 3.492


```{r}
grades <- read.csv("../data/grade_distributions.csv")

grades <- grades %>%
  mutate(
    total_points =
      a_count  * 4.0 +
      ab_count * 3.5 +
      b_count  * 3.0 +
      bc_count * 2.5 +
      c_count  * 2.0 +
      d_count  * 1.0 +
      f_count  * 0.0,
    
    total_students =
      a_count + ab_count + b_count + bc_count + c_count + d_count + f_count,
    
    gpa = total_points / total_students
  ) %>%
  drop_na() #the part that cleans the dataset

ggplot(grades, aes(x = gpa)) +
  geom_histogram(
    color = "white",
    fill = "lightblue"
  ) +
  labs(
    title = "Distribution of Course GPA",
    x = "GPA",
    y = "Count of Course Sections"
  ) +
  geom_vline(xintercept = mean(grades$gpa), size = 1, color = "red") + 
  geom_vline(xintercept = 3.492, size = 1, color = "blue")

t.test(grades$gpa, mu = 3.492, alternative = "two.sided")


```
confidence interval: 3.558, 3.563


Discussion:


